\subsection{Undefinability}
\subsubsection*{Cantor's Theorem and Cardinality Arguments}
We will show that for every infinite structure $C$ there is a subset $X\subseteq U^C$ which is \emph{not} definable over $C$. This result is a corollary to the celebrated Cantor Diagonal Theorem.
\begin{theorem}[Cantor]\label{cantordiag-thm}
Let $U$ be an infinite set and let $V_1, V_2, \ldots$ be a sequence of subsets of $U$. There is subset $W$ of $U$ such that for all $i\geq 1$, $W\neq V_i$.
\end{theorem}
\emph{Proof}: Suppose $U$ is an infinite set. Let $U^*= \{a_1, a_2, \ldots\}$ be a countably infinite subset of $U$ and let $V_1, V_2, \ldots$ be a sequence of subsets of $U$. Let $W=\{i\mid a_i\not\in V_i\}$. Note that for every $i$, $a_i\in W$ if and only if $a_i\not\in V_i$. It follows that for all $i$, $W\neq V_i$. \qed

\begin{aside}
    The idea in the above proof is to show that, regardless of which way we list the subsets $V_i$ of $U$, there will always be some other subset $W$ of $U$ which is not in the list. We construct $W$ by making sure it differs from each $V_i$ by at least one element; to do this, it suffices to let $a_i \in W$ iff $a_i \not \in V_i$. 
\end{aside}

In order to apply Theorem \ref{cantordiag-thm} to questions about definable sets we require the following result.
\begin{theorem}\label{countschema-thm}
For every structure $C$, there is a sequence $V_1,V_2,\ldots$ of subsets of $U^C$ such that for every set $X$ definable over $C$, there is an $i$ such that $X=V_i$. 

For those who know a little bit of set theory, this says that there are only countably many definable subsets of any given structure.
\end{theorem}
\emph{Proof}: Every schema is a finite sequence of symbols drawn from a finite alphabet. Thus, we may arrange all schemata $S(x)$ in a list $S_1(x), S_2(x),\ldots$, first ordered by length, and then within length, alphabetically. We obtain a list $V_1,V_2,\ldots $ of all the sets definable over $C$ by setting $V_i=S_i[C]$ for all $i$. \qed

As Theorems \ref{cantordiag-thm} and Theorem \ref{countschema-thm} entails that we can list all definable subsets of an infinite structure $C$, and theorem \ref{cantordiag-thm} entails that no list can exhaust all the definable subsets of an infinite set. So we have our result:

\begin{corollary}
For every infinite structure $C$ there is a subset $X\subseteq U^C$ which is not definable over $C$.
\end{corollary}

\subsubsection*{The Compactness Theorem and Automorphisms of ``Non-standard Models''}

Of course, this gives us no idea which particular sets are not definable over a given infinite structure. In the case of the graph $B$ introduced above, we will show that if a set is neither finite nor co-finite, it is \emph{not} definable over $B$. In order to establish this, we will deploy one of the fundamental properties of polyadic quantification theory: \emph{compactness}. First, some definitions requisite to the Compactness Theorem for Polyadic Quantification Theory.

\begin{definition}
A schema $S$ is \emph{satisfiable} if and only if for some structure $A$, $A\models S$.
\end{definition}

\begin{definition}
A set of schemata $\Gamma$ is \emph{satisfiable} if and only if there is structure $A$ such that for every schema $S\in \Gamma$, $A\models S$.
\end{definition}

\begin{definition}
A set of schemata $\Gamma$ is \emph{finitely satisfiable} if and only if for every finite set $\Delta\subseteq\Gamma$, $\Delta$ is satisfiable.
\end{definition}

\begin{theorem}[Compactness Theorem]\label{compact-thm}
For every set $\Gamma$ of schemata  of polyadic quantification theory, if $\Gamma$ is finitely satisfiable, then $\Gamma$ is satisfiable. 
\end{theorem}

Though the Compactness Theorem makes no mention of the notion of a derivation, one of its well-known proofs proceeds via the elaboration of a sound and complete formal system for logical deduction. This development will occupy our attention for much of the remainder of the course. For the moment, let's see how we can apply the Compactness Theorem to complete the analysis of the definable subsets of the structure $B$ specified above.
\begin{theorem}
If $V\subseteq U^B$ is definable over $B$, then $V$ is finite or $V$ is co-finite.
\end{theorem}
\emph{Proof}:
Suppose toward a contradiction that a schema $\sigma(x)$ defines a set $V$ which is neither finite nor co-finite over $B$. Let $\Lambda = \{ S\mid B\models S\}$; $\Lambda$ is the set of all schemata true in the structure $B$ and is often called the \emph{complete theory} of $B$. Let $y$ and $z$ be fresh variables which occur nowhere in $\Lambda$, $\sigma(x)$, or any of the schemata $S^n(x)$ for $n\geq 0$ (recall that $S^n(x)$ says that $x$ is the $n^{th}$ successor of the unique element with no predecessors). 

Define the set of schemata $\Gamma$ as follows.
\[
\Gamma = \Lambda\cup\{y\neq z \land \sigma(y) \land \neg \sigma(z)\}\cup\{\neg S^n(y) \land \neg S^n(z)\mid n\geq 0\}.
\]

Let $\Delta$ be a finite subset of $\Gamma$. As both $\sigma[B]$ and $\neg \sigma[B]$ are infinite by hypothesis, $\Delta$ can be satisfied by $B$ with suitable assignments from $U^B$ to the variables $y$ and $z$. Hence, by the Compactness Theorem, $\Gamma$ itself is satisfiable. Of course, if the structure $C$ satisfies $\Gamma$, then $C$ is not isomorphic to $B$ since the the elements of $U^C$ assigned to $y$ and $z$ in $C$ (call them $a$ and $b$ respectively) are not reachable in $C$ from the unique element of $C$ with no predecessor (whereas every element $b \in B$ is reachable in this manner). 

We will show that there is an automorphism $h$ of $C$ with $h(a)=b$. This will yield the desired  contradiction, since $C\models \sigma(y|a)$ and $C\models \neg \sigma(z|b)$. 

Note that $B$, and hence $C$, satisfy the following schemata.
\begin{itemize}
\item
$(\exists x)(\forall y)((\forall z)\neg Lzy \equiv x=y)$
\item
$(\forall x)(\exists y)(\forall z)(Lxz\equiv z=y)$
\item
$(\forall x)(\forall y)(\forall z)((Lxz\wedge Lyz)\supset x=y)$
\item
$(\forall x)\neg Lxx\\
\vdots\\
(\forall x)(\forall y_1)\ldots(\forall y_n)\neg Lxy_1\wedge Ly_1y_2\ldots\wedge Ly_nx\\
\vdots$
\end{itemize} 
The first three schemata guarantee that $L^C$ is an injective functional relation which is ``almost'' surjective -- there is a unique element of $U^C$ which lacks a pre-image under the function whose graph is $L^C$. Note that this guarantees that $U^C$ is infinite. 

\begin{aside}
    Why does this ensure that $U^C$ is infinite?
\end{aside}

The final infinite list of schemata guarantee that the the function whose graph is $L^C$ contains no finite cycles. Since $C$ is not isomorphic to $B$, all this implies that $C$ consists of an $L^C$ chain that is isomorphic to $B$ and a non-empty set of $L^C$ chains each of which is isomorphic to $\mathbb{Z}$ (the set of all integers) equipped with its usual successor relation. But, since $a$ and $b$ must lie on one or two of these ``$\mathbb{Z}$-chains,'' there is an automorphism $h$ of $C$ with $h(a)=b$ (if they lie on a single $\mathbb{Z}$-chain, shifting the $\mathbb{Z}$-chain works as an automorphism, whereas if they lie on two $\mathbb{Z}$-chains, interchanging the $\mathbb{Z}$-chains suffices). \qed